---
title: "p8105_hw2_wm2460"
author: "Wenyan Ma"
date: "2022-09-28"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```


## Problem 1: NYC Transit data

#### Importing and tidying in data
```{r q1_import_tidy}
nytransit_df = 
  read_csv(
    "./HW2_Data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(line, starts_with('station'), starts_with('route'), entry, entrance_type, vending, ada, -station_location) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

#### Short paragraph about the data

The cleaned NY transit data include variables like `r names(nytransit_df)`

After import the .csv file, my data cleaning processes include general cleaning of the variable names, selecting necessary variables into the data frame, and changing variable types.

The dimension of the cleaned NY transit data frame is `r dim(nytransit_df)`

#### Answer the questions

*   How many distinct stations are there? 
    * There are `r nrow(distinct(nytransit_df, line, station_name))` distinct stations.

```{r}
nrow(distinct(nytransit_df, line, station_name))
```

*   How many stations are ADA compliant?
    * There are `r nrow(filter(nytransit_df, ada == TRUE))` stations that are ADA compliant.
  
```{r}
nrow(filter(nytransit_df, ada == TRUE))
```

*   What proportion of station entrances / exits without vending allow entrance? 
    * The proportion of station entrances / exists without vending allow entrance is 0.377.
    
```{r proportion_calculation}
#number of observations with no vending:
no_vending =
  nrow(filter(nytransit_df, vending == 'NO'))
no_vending

#number of observations with entry but no vending: 
entry_true = 
  nrow(filter(nytransit_df, vending == 'NO', entry == 'TRUE'))
entry_true

#proportion of station entrances / exits without vending allow entrance
proportion = entry_true/no_vending 
proportion

```


#### Reformat route name and route number
```{r}
nytransit_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

nytransit_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
  
  
## Problem 2: Trash Wheel

#### Read and clean the Mr. Trash Wheel sheet

```{r mr_trash_wheel}
mrtrash_df = 
  read_excel("./HW2_Data/Trash Wheel Collection Data.xlsx",
  sheet = "Mr. Trash Wheel",
  range = "A2:N549") %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = round(sports_balls), 
         sports_balls = as.integer(sports_balls))
```

#### Read and clean the Professor Trash Wheel sheet

```{r prof_trash_wheel}
proftrash_df = 
  read_excel("./HW2_Data/Trash Wheel Collection Data.xlsx",
  sheet = "Professor Trash Wheel",
  range = "A2:M96") %>% 
  janitor::clean_names() 
```


#### Combine the two 
```{r combine}
mrtrash_df$dumpster = as.numeric(mrtrash_df$dumpster)
mrtrash_df$year = as.numeric(mrtrash_df$year)


# Add a variable for each dataset to distinguish the original dataset name
mrtrash_df$dataset = "Mr"
proftrash_df$dataset = "Prof"

bind_rows(mrtrash_df, proftrash_df)

#Combine the two
join_trash_df = 
  full_join(mrtrash_df, proftrash_df)
```

#### Description of the joined dataset

  The joined dataset combined the "Mr. Trash Wheel" and "Professor Trash Wheel" sheets. 
  There are `r nrow(join_trash_df)` observations and `r ncol(join_trash_df)` variables in the joined dataset. 
  Some key variables include `weight_tons`, `volume_cubic_yards`, and `date`. 
  The total weight of trash collected by Professor Trash Wheel is `r proftrash_df %>% pull(weight_tons) %>% sum` tons.
  The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r mrtrash_df %>% filter(year == 2020) %>% pull(sports_balls) %>% sum`.
    
    
## Problem 3: FiveThirtyEight

#### Clean pols-month.csv
```{r pols_month}
pols_month = 
  read_csv("./HW2_Data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(month = as.integer(month),
         month = month.name[month],
         year = as.integer(year),
         day = as.integer(day),
         president = case_when(prez_dem == "1" ~ "dem", 
                               prez_gop == "1" ~ "gop")) %>% 
  arrange(year, month) %>% 
  select(-day, -prez_dem, -prez_gop)
  
```


#### Clean snp.csv
```{r snp}
snp = 
  read_csv("./HW2_Data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year","month", "day"), convert = TRUE) %>%
  mutate(month = as.integer(month),
         month = month.name[month],
         year = as.integer(year),
         day = as.integer(day)) %>% 
  select(-day) %>% 
  arrange(year, month) %>% 
  relocate(year, month, close)
```


#### Clean unemployment.csv
```{r unemployment}
unemployment = 
  read_csv("./HW2_Data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment_rate") %>% 
  mutate(year = as.numeric(year),
         month = recode(month,
                   jan = "Janurary",
                   feb = "Feburary", 
                   mar = "March", 
                   apr = "April",
                   may = "May",
                   jun = "June",
                   jul = "July",
                   aug = "August",
                   sep = "September",
                   oct = "October",
                   nov = "November",
                   dec = "December")) %>% 
  na.omit()
```

#### Merging

```{r merging}
join = 
  left_join(pols_month, snp, by = c("year", "month")) %>% 
  left_join(unemployment, by = c("year", "month"))
```

#### Description about the dataset

  In the `pols_month` dataset, there are `r nrow(pols_month)` observations and `r ncol(pols_month)` variables. The variables in this dataset include `r colnames(pols_month)`.
  
  In the `snp` dataset, there are `r nrow(snp)` observations and `r ncol(snp)` variables. The variables in this dataset include `r colnames(snp)`.
  
  In the `unemployment` dataset, there are `r nrow(unemployment)` observations and `r ncol(unemployment)` variables. The variables in this dataset include `r colnames(unemployment)`.

  The resulting dataset, `join`, contains `r nrow(join)` observations and `r ncol(join)` variables. The dimension of the resulting dataset is `r dim(join)`. The range of years is between `r min(join$year)` and `r max(join$year)`. Some key variables in the dataset include `president`, `unemployment_rate`, and `close`. 



#### End of HW2



